---
title: "Configuration"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{a01_configuration}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Initiating database connection

We use the **CDMConnector** package to establish the database connection. You can configure the connection either by reading credentials from a `.Renviron` file or explicitly writing them in your script.

```{r, echo=TRUE, eval=FALSE}
################################################################################
#
# Initiate the database connection
#
#################################################################################

user <- Sys.getenv("DB_USERNAME")
pw <- Sys.getenv("DB_PASSWORD")
server <- stringr::str_c(Sys.getenv("DB_HOST"), "/", Sys.getenv("DB_NAME"))
port <- Sys.getenv("DB_PORT")

cdmSchema <-
  Sys.getenv("OHDSI_CDM") # Schema which contains the OHDSI Common Data Model
cdmVocabSchema <-
  Sys.getenv("OHDSI_VOCAB") # Schema which contains the OHDSI Common Data Model vocabulary tables.
cdmResultsSchema <-
  Sys.getenv("OHDSI_RESULTS") # Schema which contains "cohort" table (is not mandatory)
writeSchema <-
  Sys.getenv("OHDSI_WRITE") # Schema for temporary tables, will be deleted
writePrefix <- "c2t_"

db = DBI::dbConnect(
  RPostgres::Postgres(),
  dbname = Sys.getenv("DB_NAME"),
  host = Sys.getenv("DB_HOST"),
  user = Sys.getenv("DB_USERNAME"),
  password = Sys.getenv("DB_PASSWORD"),
  port  = port
)

cdm <- CDMConnector::cdmFromCon(
  con = db,
  cdmSchema = cdmSchema,
  achillesSchema = cdmResultsSchema,
  writeSchema = c(schema = writeSchema, prefix = writePrefix),
)
```

## Configuration

To enhance the user experience, an R environment named `studyEnv` is created, which consolidates the necessary parameters required for running the study. This environment simplifies managing inputs and ensures consistency across functions. Input description available [here](../reference/cohort2TrajectoryConfiguration.html). Below is an example configuration with default values: 

```{r, echo=TRUE, eval=FALSE}
################################################################################
#
# Configure study settings
#
################################################################################

studyEnv <- cohort2TrajectoryConfiguration(
  cdm = cdm,
  studyName = "Cohort2Trajectory",
  baseUrl = "http://localhost:8080/WebAPI",
  pathToStudy = getwd(),
  atlasTargetCohort = NULL,
  atlasStateCohorts = NULL,
  stateCohortLabels = NULL,
  stateCohortPriorityOrder = NULL,
  stateCohortMandatory = NULL,
  stateCohortAbsorbing = NULL,
  stateSelectionType = NULL,
  oocFix = "None",
  trajectoryType = "Discrete",
  lengthOfStay = NULL,
  outOfCohortAllowed = FALSE,
  runSavedStudy = FALSE,
  useCDM = TRUE,
  trajectoryDataObject = NULL,
  pathToData = './Cohort2Trajectory/Data/importedData.csv',
  allowedStatesList = createStateList(stateCohortLabels),
  mergeStates = FALSE,
  mergeThreshold = 0.5,
  batchSize = 1000
) 
```

<br>

### Configuration with ATLAS generated cohorts

```{r, echo = TRUE, eval = FALSE}

studyEnv <- cohort2TrajectoryConfiguration(
    baseUrl = "http://localhost:8080/WebAPI",     #<--- URL for API
    useCDM = TRUE,                                #<--- set use CDM TRUE
    studyName = "nameForStudy",
    pathToStudy = getwd(),
    atlasTargetCohort = 1,                 #<--- existing cohort's id
    atlasStateCohorts = c(2, 3),           #<--- existing cohorts' ids
    stateCohortLabels = c("label_for_cohort1", "label_for_cohort2"),
    stateCohortMandatory = c("label_for_cohort2"),
    stateCohortAbsorbing = c("label_for_cohort2"),
    outOfCohortAllowed = FALSE,
    trajectoryType = "Discrete",
    lengthOfStay = 30,
    stateSelectionType = "Priority",
    stateCohortPriorityOrder = c("label_for_cohort1", "label_for_cohort2"),
    runSavedStudy = FALSE,
    batchSize = 10
  )

```

<br>

### Configuration with saved study 

```{r, echo = TRUE, eval = FALSE}

studyEnv <- cohort2TrajectoryConfiguration(
    baseUrl = NULL,     
    useCDM = FALSE,                                
    studyName = "nameForStudy",                   #<--- name of the study in trajectorySettings.csv
    pathToStudy = getwd(),
    stateCohortLabels = c("label_for_cohort1", "label_for_cohort2"),
    runSavedStudy = TRUE,                           #<--- set run saved study to true
    batchSize = 10
  )

```

<br>

### Configuration with data as input

```{r, echo = TRUE, eval = FALSE}

studyEnv <- cohort2TrajectoryConfiguration(
    baseUrl = NULL,   
    useCDM = FALSE,                       
    trajectoryDataObject = rawData,               #<--- eg output of read.csv
    studyName = "nameForStudy",
    pathToStudy = getwd(),
    stateCohortMandatory = c("label_for_cohort2"),
    stateCohortAbsorbing = c("label_for_cohort2"),
    outOfCohortAllowed = FALSE,
    trajectoryType = "Discrete",
    lengthOfStay = 30,
    stateSelectionType = "Priority",
    stateCohortPriorityOrder = c("label_for_cohort1", "label_for_cohort2"),
    runSavedStudy = FALSE,
    batchSize = 10
  )

```

<br>

### Configuration with path to data as input

```{r, echo = TRUE, eval = FALSE}

studyEnv <- cohort2TrajectoryConfiguration(
    baseUrl = NULL,   
    useCDM = FALSE,                       
    pathToData = paste0(getwd(), "/path/to/data/data.csv")), #<--- path to data as input
    studyName = "nameForStudy",
    pathToStudy = getwd(),
    stateCohortMandatory = c("label_for_cohort2"),
    stateCohortAbsorbing = c("label_for_cohort2"),
    outOfCohortAllowed = FALSE,
    trajectoryType = "Discrete",
    lengthOfStay = 30,
    stateSelectionType = "Priority",
    stateCohortPriorityOrder = c("label_for_cohort1", "label_for_cohort2"),
    runSavedStudy = FALSE,
    batchSize = 10
  )

```
