---
title: "Import data"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{a02_import}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Importing data for study 

This section provides a detailed explanation of how to use `getDataForStudy` function. For a general overview, refer to the [Getting Started](a00_introduction.html#study-configuration) section. Regarding data sources, there are three options: providing existing data, running a saved study, or using the CDM as a data source. 


```{r, eval = FALSE, echo = TRUE}
getDataForStudy(cdm = cdm, studyEnv = studyEnv)
```


If the user opts out of using the previous configuration, the necessary inputs must be provided manually. Below is a sample with default values: 

```{r, eval = FALSE, echo = TRUE}
getDataForStudy(
  runSavedStudy = studyEnv$runSavedStudy,
  useCDM = studyEnv$useCDM,
  trajectoryDataObject = NULL,
  cdm = NULL,
  studyEnv = NULL,
  atlasStateCohorts = studyEnv$atlasStateCohorts,
  atlasTargetCohort = studyEnv$atlasTargetCohort,
  stateCohortLabels = studyEnv$stateCohortLabels,
  baseUrl = studyEnv$baseUrl,
  outOfCohortAllowed = studyEnv$outOfCohortAllowed,
  stateCohortMandatory = studyEnv$stateCohortMandatory,
  stateCohortPriorityOrder = studyEnv$stateCohortPriorityOrder,
  mergeStates = studyEnv$mergeStates,
  mergeThreshold = studyEnv$mergeThreshold,
  pathToData = NULL
)
```

<br>

## Use CDM

To use the CDM as a data source, set the `useCDM` variable to `TRUE`. The cohorts specified in `atlasStateCohorts` and `atlasTargetCohort` will be used. It is recommended to set `trajectoryDataObject`, `pathToData`, and `runSavedStudy` to `NULL`.

```{r, echo = TRUE, eval = FALSE}
getDataForStudy(
  runSavedStudy = NULL,
  useCDM = studyEnv$useCDM,
  trajectoryDataObject = NULL,
  cdm = NULL,
  studyEnv = NULL,
  atlasStateCohorts = studyEnv$atlasStateCohorts,
  atlasTargetCohort = studyEnv$atlasTargetCohort,
  stateCohortLabels = studyEnv$stateCohortLabels,
  baseUrl = studyEnv$baseUrl,
  outOfCohortAllowed = studyEnv$outOfCohortAllowed,
  stateCohortMandatory = studyEnv$stateCohortMandatory,
  stateCohortPriorityOrder = studyEnv$stateCohortPriorityOrder,
  mergeStates = studyEnv$mergeStates,
  mergeThreshold = studyEnv$mergeThreshold,
  pathToData = NULL
)
```

<br>

### Without atlas support 

To use the CDM without ATLAS support, ensure the required JSON files are located under `studyName/JSON`. In this case, set the `baseUrl` variable to `NULL`.

```{r, echo = TRUE, eval = FALSE}
getDataForStudy(
  runSavedStudy = NULL,
  useCDM = studyEnv$useCDM,
  trajectoryDataObject = NULL,
  cdm = NULL,
  studyEnv = NULL,
  atlasStateCohorts = studyEnv$atlasStateCohorts,
  atlasTargetCohort = studyEnv$atlasTargetCohort,
  stateCohortLabels = studyEnv$stateCohortLabels,
  baseUrl = NULL,
  outOfCohortAllowed = studyEnv$outOfCohortAllowed,
  stateCohortMandatory = studyEnv$stateCohortMandatory,
  stateCohortPriorityOrder = studyEnv$stateCohortPriorityOrder,
  mergeStates = studyEnv$mergeStates,
  mergeThreshold = studyEnv$mergeThreshold,
  pathToData = NULL
)
```

<br>

## Run saved study

If a saved study configuration is used, the previously set options are overridden by the saved settings. It is recommended to set `trajectoryDataObject`, `pathToData`, and `useCDM` to `NULL`.

```{r, echo = TRUE, eval = FALSE}
getDataForStudy(
  runSavedStudy = studyEnv$runSavedStudy,
  useCDM = NULL,
  trajectoryDataObject = NULL,
  cdm = NULL,
  studyEnv = NULL,
  atlasStateCohorts = studyEnv$atlasStateCohorts,
  atlasTargetCohort = studyEnv$atlasTargetCohort,
  stateCohortLabels = studyEnv$stateCohortLabels,
  baseUrl = studyEnv$baseUrl,
  outOfCohortAllowed = studyEnv$outOfCohortAllowed,
  stateCohortMandatory = studyEnv$stateCohortMandatory,
  stateCohortPriorityOrder = studyEnv$stateCohortPriorityOrder,
  mergeStates = studyEnv$mergeStates,
  mergeThreshold = studyEnv$mergeThreshold,
  pathToData = NULL
)
```

<br>

## Add data object or path to data 

To add data manually, provide either the `trajectoryDataObject` or the `pathToData` as input. It is suggested to set both `useCDM` and `runSavedStudy` to `NULL`.

For appending personal data to existing datasets, refer to the [create trajectories](a03_trajectories.html#adding-personal-data-to-self-inserted-data) section.


