---
title: "Create trajectories"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{a03_trajectories}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Creating trajectories

This section outlines the process for creating trajectories, including handling edge cases.

The standard approach involves using the default configuration, data import, and pre-processing steps, as described in [getting started](a00_introduction.html#study-configuration).

```{r, eval = FALSE, echo = TRUE}
createTrajectories (cdm = cdm, studyEnv = studyEnv)
```

If the user opts out of using the previous configuration, the necessary inputs must be provided manually. Below is list of default values:

```{r, eval = FALSE, echo = TRUE}
createTrajectories(
  cdm = NULL,
  studyEnv = NULL,
  trajectoryType = studyEnv$trajectoryType,
  runSavedStudy = studyEnv$runSavedStudy,
  oocFix = studyEnv$oocFix,
  outOfCohortAllowed = studyEnv$outOfCohortAllowed,
  lengthOfStay = studyEnv$lengthOfStay,
  stateCohortLabels = studyEnv$stateCohortLabels,
  stateCohortPriorityOrder = studyEnv$stateCohortPriorityOrder,
  stateSelectionType = studyEnv$stateSelectionType,
  stateCohortAbsorbing = studyEnv$stateCohortAbsorbing,
  stateCohortMandatory = studyEnv$stateCohortMandatory,
  allowedStatesList = studyEnv$allowedStatesList,
  useCDM = studyEnv$useCDM,
  pathToStudy = studyEnv$pathToStudy,
  batchSize = studyEnv$batchSize
)
```

## Adding personal data to self-inserted data

To include personal data such as patient age and gender while creating trajectories, ensure `useCDM` is set to `TRUE`. 

```{r, eval = FALSE, echo = TRUE}
createTrajectories (
  cdm = cdm,
  studyEnv = studyEnv,
  useCDM = TRUE
)
```

The function returns a dataframe with the trajectories and saves the trajectories to the study path.